---
title: "Data"
format:
  html:
    toc: true
    toc-depth: 6
    toc-title: Contents
    toc-location: left
    toc-expand: 5
    smooth-scroll: true
    anchor-sections: true
    include-after-body: abbrv_toc.html
    number-sections: true
---

## General Information About the Data

*This part is going to be updated soon...*

**Data source:** [Türkiye Health Survey 2022](https://data.tuik.gov.tr/Bulten/Index?p=Turkiye-Saglik-Arastirmasi-2022-49747)

### Why We Selected This Data

*This part is going to be updated soon...*

### Our Objectives

*This part is going to be updated soon...*

## Importing and Preprocessing Data: Our Approach

Before importing our data, we manually cleaned it by removing Turkish text and unnecessary information. Then, we used the '*dplyr*' and '*tidyr*' packages to improve our understanding of the data. The relevant code is provided below.

This study includes three datasets:

### **Percentage of Health Problems in the Last 12 Months by Sex, 2016-2022** {toc-text="Dataset 1"}

```{r}
#| code-fold: true
#| code-summary: "Show the code"

```

### **Percentage of Status of Alcohol Use by Sex and Age Group, 2016-2022** {toc-text="Dataset 2"}

```{r}
#| code-fold: true
#| code-summary: "Show the code"
suppressPackageStartupMessages(library(dplyr))
library(readxl)
library(dplyr)
library(tidyr)

data_2 <- read_excel("tidydataset2.xls")
colnames(data_2) <- c("age", "men_2016", "women_2016", "men_2019", "women_2019", "men_2022", "women_2022", "usage")
data_2_longer <- data_2 %>%
  pivot_longer(cols = starts_with(c("men", "women")),
               names_to = "gender",
               values_to = "rate") %>%
  separate(gender, into = c("gender", "year"), sep = "_") %>%
  arrange(age) %>%
  mutate_at(vars(year, rate), as.numeric) %>%
  na.omit()
```

### **Body Mass Index Distribution of Individuals by Sex, 2008-2022** {toc-text="Dataset 3"}

```{r}
#| code-fold: true
#| code-summary: "Show the code"
suppressPackageStartupMessages(library(dplyr))
library(readxl)
library(dplyr)
library(tidyr)

data_3 <- read_excel("tidydataset3.xls")
colnames(data_3) <- c("Year", "Sex", "Underweight", "Normal_weight", "Pre_Obese", "Obese")
data_3_long <- data_3 %>%
  pivot_longer(cols = c(Underweight, Normal_weight, Pre_Obese, Obese),
               names_to = "Category",
               values_to = "Percentage") %>%
  mutate(sex_group = ifelse(Sex == "Total", "Total", "Individual")) %>%
  group_by(sex_group) %>%
  ungroup() %>%
  dplyr::filter(Sex != "Total" | n() == 1) %>%
  select(-sex_group)
```

## Exploratory Data Analysis

*This part is going to be updated soon...*
